\mutagen: Coverage-Guided, Property-Based Testing using Exhaustive Structure-Preserving Mutations

### 2022-03-08 papers/fse22/main.tex
Added formal definition of mutations. Other fixes.

@@

Unlike the generic bit-level mutators often used in the fuzzing community
\cite{}, structure-preserving mutations specified at the data type level can
effectively produce syntactically valid mutants. 
%
% %
% High-level mutators avoid wasting time with generic low-level mutators that work
% at the bit level and know very little about data structures, thus producing
% syntactically broken mutants most of the time.
%
This grammar-aware technique has shown to be effective when fuzzing systems
accepting structurally complex inputs \cite{holler2012fuzzing, wang2019superion,
xsmith}.
%
In practice, most tools use existing grammars to tailor mutations to the input
structure used by the system under test. 
%
% Notably external grammars are not required in CGPT.
%
%% Tools like LangFuzz \cite{holler2012fuzzing}, Superion \cite{wang2019superion},
%% XSmith \cite{xsmith} use existing grammars to tailor the mutators to the
%% specific input structure used by the system under test.
%
% In CGPT, external grammars are not required.
%
In contast, CGPT relies on the data type information of the inputs to the
testing properties to derive specialized structure-preserving mutators directly
--- making strongly-typed programming languages an ideal match for this
technique.
%
% already provide a good source of information about how data can be mutated,
% e.g., by changing one data constructor by another.
%
%
% The data types used by the inputs of the testing properties already describe the
% structure of the random data one needs to mutate in a concrete manner, and
% Specialized mutators can be automatically derived directly from data types
% definitions ---
%

@@

%% Notably, \emph{scheduling mutations exhaustively does not mean that mutations
%%   are always exhaustive.}
%
Internally, our tool distinguishes two kinds of mutations.
%  those that can be computed
% deterministically, 
% and those that can be
% obtained non-deterministically.
%
On one hand, \emph{deterministic (pure) mutations} encode transformations that
yield a single mutated test case obtained by swapping data constructors around,
as well as rearranging or returning subexpressions.
%
On the other hand, \emph{non-deterministic (random) mutations} are used to
represent transformations over large enumeration types.
%
% like numbers or characters.
%
This mechanism let us selectively escape the scalability issues of
exhaustiveness by yielding a random generator that replaces a specific
subexpression of an input test case with a randomly generated one.
%
Such generator is then sampled a \emph{small} number of times. 
%
This way \mutagen avoids, for instance, mutating every numeric subexpresion of a
test case into every other possible number.

@@

The second heuristic adjusts the number of test cases generated from random
mutations  
%
% number of times our tool samples random
% mutants, i.e., those corresponding to large enumeration types as described
% above.
%
by monitoring how often we generate interesting test cases.
%
If this frequency stalls, \mutagen resets the testing loop and increases the
effort put into sampling random mutations.
%
This automatically finds a suitable value for this parameter on the fly.

@@

%% Last two sections
Finally, we discuss related work in \S \ref{sec:related} and conclude
in \S \ref{sec:conclusions}.

@@

\item \emph{Power schedule:}
%
it is unclear how well the heuristic that assigns a budget to each interesting
test case in CGPT performs in tandem with structure-preserving mutations derived
for an arbitrary data type.
%
On one hand, spending too much budget on not-so-interesting test cases might
prevent bugs from being discovered on a timely basis.
%
In turn, spending too little budget on interesting test cases might prevent bugs
from being discovered at all --- randomly generating the same initial test case
again later does not help, as it will not be considered interesting again
according to historic trace information.
%
Finding a well behaved heuristic between these two scenario can be quite
challenging in the general case. 
%
\emph{Ideally, the scheduling mechanism should be as unbiased as possible.}

@@

In \mutagen, mutations are defined as the set of mutants that can be obtained by
transforming the input test case \emph{at the top-level} (the root node).
%
We define the mutations of a value of type \texttt{a} (written \texttt{Mutations
  a}) as \emph{a function that takes a value of such a type and returns a list
  of mutants}.
% (written \texttt{Mutant a}).
%
In Haskell, we introduce the type synonym:
%
\begin{verbatim}
type Mutations a = a -> [Mutant a]
\end{verbatim}
%
\noindent As mentioned earlier, concrete mutants can be obtained either from a
pure or a random mutation:

@@

\paragraph{Pure mutants} 
% are used most of the time, and 
%
encode simple deterministic transformations over the outermost data constructor
of the input --- recursive mutations will be introduced soon.
%
These transformations can:
%
\begin{inparaenum}
\item return an immediate subexpression of the same type as the input;
\item swap the outermost data constructor with any other constructor of the
  same type, reusing existing subexpressions; and
\item rearrange and replace fields using existing ones of the same type.
\end{inparaenum}

\setlength{\intextsep}{10pt}%
\setlength{\columnsep}{10pt}%
\begin{wrapfigure}{r}{0.2\textwidth}
\vspace{-10pt}
\begin{verbatim}
mutate t = case t of 
  Leaf x -> [ 
    PURE (Branch def x def)
  ]
  Branch l x r -> [ 
    PURE l, PURE r, 
    PURE (Leaf x), 
    PURE (Branch l x l),
    PURE (Branch r x r),
    PURE (Branch r x l)
  ]
\end{verbatim}
\vspace{-10pt}
\caption{\label{fig:mutagen:mutator}\mutagen mutator for the \texttt{Tree} data
  type.}
\vspace{-10pt}
\end{wrapfigure}

@@

@@

%% At this point, the reader might consider:
%% %
%% what about large enumeration types like integers or characters?
%% %
%% Does \mutagen transform them exaustively too?
%% %
%% \emph{Certainly not.}
%
\paragraph{Random mutants} let us \emph{selectively avoid exhaustiveness} when
  mutating values of large enumeration types (e.g. numbers).
  % 
  % --- or any
  % other type the user might want to use random mutations with.
%
%
Instead of creating a new \texttt{PURE} mutant for every other numerical
subexpression exhaustively, we condense them into a generator that can be
repeatedly sampled to produce new random values.   
%
% small number of values from a
% given random generator.
%
This way, a mutator for integers simply becomes:

@@

So far we have defined mutations that transform only the root of the input.
%
To apply these mutations to every internal subexpression we use two utility
functions that can be automatically synthesized.
%
Firstly, a function \texttt{Positions} traverses the input and builds a Rose
tree\cite{} of \emph{mutable positions}, where positions are lists of indices
encoding the path from the root to every mutable subexpression.
%
% For instance, the positions of a \texttt{Tree} are computed as follows:
%
% \begin{verbatim}
% positions (Leaf x) =
%   node [ (0, positions x) ]
% positions (Branch l x r) =
%   node [ (0, positions l), (1, positions x), (2, positions r) ]
% \end{verbatim}
%
% \noindent where the function \texttt{node} simply builds a node of the positions
% Rose tree and propagates the accumulated prefix (the path from the root) to its
% children.
%
For instance, the mutable positions of the value \texttt{Branch (Leaf 1) 2
  (Leaf 3)} are as follows:

\vspace{-5pt}
\begin{equation}
  \nonumber
  \footnotesize
  \texttt{Positions}\quad
  \left(
  \tikz[
    baseline=-6mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\texttt{Branch}}
      child {node {\texttt{Leaf}}
        child {node {\texttt{1}}}
      }
      child {node {\texttt{2}}}
      child {node {\texttt{Leaf}}
        child {node {\texttt{3}}}
      };
  }
  \right)
  \quad
  =
  \quad
  \tikz[
    baseline=-6mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\texttt{[]}}
      child {node {\texttt{[0]}}
        child {node {\texttt{[0,0]}}}
      }
      child {node {\texttt{[1]}}}
      child {node {\texttt{[2]}}
        child {node {\texttt{[2,0]}}}
      };
  }
\end{equation}

Then, we define a function \texttt{MutateInside} that takes a desired position
within an input test case and mutates its corresponding subexpression, returning
a list of mutants.
%
%% given the desired position to mutate within a test case, we
%% define another function \texttt{inside} that finds the subexpression
%% corresponding to it and mutates it.
%
% A slightly simplified version of this function for the \texttt{Tree} data type is
% as follows:
%
% \begin{verbatim}
% inside []     t              = mutate t
% inside (0:ps) (Leaf x)       = [ Leaf x'       | x' <- inside ps x ]
% inside (0:ps) (Branch l x r) = [ Branch l' x r | l' <- inside ps l ]
% inside (1:ps) (Branch l x r) = [ Branch l x' r | x' <- inside ps x ]
% inside (2:ps) (Branch l x r) = [ Branch l x r' | r' <- inside ps r ]
% \end{verbatim}
%
This function traverses the desired position, calling itself recursively until
it reaches the desired subexpression, where a mutation encoded by
\texttt{mutate} can be applied directly.
% locally at the top-level (case \texttt{inside [] t}).
%
% The rest of the \texttt{MutateInside}'s functionality simply unwraps and
% rewraps the intermediate subexpressions and is not particularly relevant for
% the point being made.

@@

\subsection{Detecting Trace Space Saturation And Tuning Random Mutations Parameter}

@@

\noindent Where $l_{pc}$, $l_{p}$, $l_{v}$, and $l_{v'}$ represent the labels
of: the program counter, the memory pointer, and the old and new values stored in
that memory cell.
%
The symbol $\vee$ simply denotes the join of two labels, i.e., the
\emph{maximum} of their sensibilities.
%
Later, bugs are systematically introduced in the IFC enforcing mechanism by
removing or weakening the checks stored in this rule table.

@@

Having sorted these issues out, we mechanically injected 10 new bugs in the
validator as well as 5 new bugs in the interpreter of this engine (see Table
\ref{table:wasm:injectedbugs}).
%
These bugs either
%
\begin{inparaenum}
\item remove an existing integrity check (to weaken the WebAssembly
  type-system/validator); or
\item simulate a copy-paste bug \cite{CPMiner},%
  \footnote{This kind of bugs was inspired by the real bug \#1 we found prior
    this step.}
  %
  replacing the implementation of an instruction with a compatible one
  (e.g., \texttt{i32.add} by \texttt{i32.sub}).
\end{inparaenum}

@@

%% We begin by designing a property to test the WebAssembly validator implemented
%% in \textit{haskell-wasm}.
%% %
Our testing approach is to assert that, whenever a randomly generated (or
mutated) WebAssembly module is valid according to \textit{haskell-wasm}, then
the reference implementation agrees upon it.
%
%% In other words, we are testing for false negatives.
%
In Haskell, we write the property:

@@

\noindent The precondition (\texttt{isValidHaskellWasm m}) runs the input
WebAssembly module \texttt{m} against \textit{haskell-wasm}'s validator, whereas
the postcondition (\texttt{isValidRefImpl m}) serializes \texttt{m}, runs it
against the reference implementation and checks that no errors are produced.

@@

\begin{verbatim}
prop_interpreter ty fun args =
  do let m = mk_module fun ty "f"
     resHs   <- invokeHaskellWasm m "f" args
     resSpec <- invokeRefImpl     m "f" args
     return (equivalent resHs resSpec)
\end{verbatim}

@@

%% If we only consider the succesfull runs where \fuzzchick does not time out, we
After carefully analyzing the results obtained using \fuzzchick, we observed a
peculiarity:
%
this tool either finds bugs quickly (after a few thousands tests) or does not
find them at all within the one-hour budget.
%% , suggesting that its power
%% scheduler assigns some enqueued test cases too little energy before they become
%% uninteresting and get ultimately thrown away.
%
With this in mind, it is fair to think that \fuzzchick might still be a better
choice if we set a shorter timeout and run it several times in a row.
%
Thus, to improve upon \fuzzchick, our tool must find bugs not only more
reliably, but also fast.

\setlength{\intextsep}{10pt}%
\setlength{\columnsep}{10pt}%
\begin{wrapfigure}{r}{0.5\linewidth}
%% \begin{wrapfigure}{r}{0.5\textwidth}
%% \centering
\vspace{-10pt}
\scriptsize

@@

\captionof{table}{\label{table:ifc:tests} Mean tests until first failure and effect size
  comparison between \fuzzchick and \mutagen when considering only successful
  runs.}
\vspace{-20pt}
%% \end{wrapfigure}
\end{wrapfigure}

Moreover, disabling the trace saturation heuristic (\textit{no reset}) we
observed two effects:
%
fixing the number of random mutations to 25 adds a constant overhead when
finding most of the (easier) bugs, suggesting that we are spending too time
mutating numeric subexpressions inside the generated machine states.
%
%% and that a
%% smaller number of random mutations could be equally effective to find such bugs.
%
However, some of the hardest-to-find bugs cannot be reliably exposed using this
fixed amount of random mutations (\#4, \#9, \#12, \#18, and \#10), suggesting
that one should increase this number even further to reliably discover all bugs
within the time budget.
%
In consequence, \emph{we argue that our heuristic is effective at automatically
tuning this parameter}.
%
%% , especially when the user is unsure about what the best value for it might be.

@@

% Limited evaluation
Our experiments cover generating test cases in two substantially different
scenarios (valid IFC machine states and WebAssembly programs). 
%
Nonetheless, we cannot generalize that \mutagen will be effective at finding
bugs in other systems or testing properties. 
%
However, our tool can automatically synthesize all the boilerplate code required
to operate, becoming a drop-in replacement of \quickcheck (the most ubiquitous
RPBT tool in Haskell). 
%     
This makes it an appealing alternative whenever \quickcheck is unable to
penetrate some sparse precondition and simply gives up.

% Coq -> Haskell translation
As mentioned in \S \ref{sec:casestudies}, our evaluation required us to
translate the IFC state machine used by \citeauthor{lampropoulos2019coverage}
from Coq to Haskell. 
%
Although this translation could be done quite mechanically (this process was
partially aided by Coq's code extraction mechanism), we cannot guarantee that
our Haskell version of the code behaves exactly as the original one. 
%
% In addition, one should also consider some  
% that the specific performance the
% runtime systems of these two programming languages can affect, the
% draw strong conclusions based only on time-to-failure.

>>>>

\newcommand\mydots{\hbox to 0.75em{.\hss.\hss.}}
\makeatletter
\newcommand{\vast}{\bBigg@{4}}
\newcommand{\Vast}{\bBigg@{5}}
\makeatother

@@

Unlike the generic bit-level mutators often used in the fuzzing community
\cite{zzuf,radamsa}, structure-preserving mutations specified at the data type
level can effectively produce syntactically valid mutants. 
%
% %
% High-level mutators avoid wasting time with generic low-level mutators that work
% at the bit level and know very little about data structures, thus producing
% syntactically broken mutants most of the time.
%
This grammar-aware technique has shown to be effective when fuzzing systems
accepting structurally complex inputs \cite{holler2012fuzzing, wang2019superion,
xsmith}.
%
In practice, most tools use existing grammars to tailor mutations to the input
structure used by the system under test. 
%
% Notably external grammars are not required in CGPT.
%
%% Tools like LangFuzz \cite{holler2012fuzzing}, Superion \cite{wang2019superion},
%% XSmith \cite{xsmith} use existing grammars to tailor the mutators to the
%% specific input structure used by the system under test.
%
% In CGPT, external grammars are not required.
%
In contast, CGPT relies on the data type information of the inputs to the
testing properties to derive specialized structure-preserving mutators directly
--- making strongly-typed programming languages an ideal match for this
technique.
%
% already provide a good source of information about how data can be mutated,
% e.g., by changing one data constructor by another.
%
%
% The data types used by the inputs of the testing properties already describe the
% structure of the random data one needs to mutate in a concrete manner, and
% Specialized mutators can be automatically derived directly from data types
% definitions ---
%

@@

%% Notably, \emph{scheduling mutations exhaustively does not mean that mutations
%%   are always exhaustive.}
%
Internally, our tool distinguishes two kinds of mutations.
%  those that can be computed
% deterministically, 
% and those that can be
% obtained non-deterministically.
%
On one hand, \emph{deterministic (pure) mutations} encode transformations that
yield a single mutated test case obtained by swapping data constructors around,
as well as rearranging or returning subexpressions.
%
On the other hand, \emph{non-deterministic (random) mutations} are used to
represent transformations over large enumeration types.
%
% like numbers or characters.
%
This mechanism let us selectively escape the scalability issues of
exhaustiveness by yielding a random generator that replaces a specific
subexpression of an input test case with a randomly generated one.
%
Such generator is later sampled a \emph{relatively small} number of times. 
%
This way \mutagen avoids, for instance, mutating every numeric subexpresion of a
test case into every other possible number.

@@

The second heuristic adjusts the number of test cases generated from random
mutations  
%
% number of times our tool samples random
% mutants, i.e., those corresponding to large enumeration types as described
% above.
%
by monitoring how often we generate interesting test cases.
%
Whenever this frequency stalls, \mutagen resets the testing loop and increases the
effort put into sampling random mutations.
%
This way, our tools automatically finds a suitable value for this parameter on
the fly.

@@

%% Last two sections
We additonally present threats to validity in \S \ref{sec:threats} and discuss
related work in \S \ref{sec:related}, to finally conclude in \S
\ref{sec:conclusions}.

@@

\item \emph{Power schedule:}
%
it is unclear how well the heuristic that assigns a budget to each interesting
test case in CGPT performs in tandem with structure-preserving mutations derived
for an arbitrary data type.
%
On one hand, spending too much budget on not-so-interesting test cases might
prevent bugs from being discovered on a timely basis.
%
In turn, spending too little budget on interesting test cases might prevent bugs
from being discovered at all --- randomly generating the same initial test case
again later does not help, as it will no longer be considered interesting
according to historic trace information.
%
Finding a well behaved heuristic between these two scenario can be quite
challenging in the general case. 
%
\emph{Ideally, the scheduling mechanism should be as unbiased as possible.}

@@

In \mutagen, mutators are defined as the set of mutants that can be obtained by
transforming the input test case \emph{at the top-level} (the root data
constructor).
%
We define mutators of a value of type \texttt{a} (written \texttt{Mutator a}) as
\emph{a function that takes a value of such a type and returns a list of
mutants}.
% (written \texttt{Mutant a}).
%
In Haskell, we introduce the type synonym:
%
\begin{verbatim}
type Mutator a = a -> [Mutant a]
\end{verbatim}
%
\noindent As mentioned earlier, concrete mutants can be obtained either from a
pure or a random mutation, which we define as follows:

@@

We first focus on pure mutants, which 
% are used most of the time, and 
%
encode deterministic transformations over the \emph{outermost} data constructor
of the input --- recursive mutations will be introduced soon.
%
These transformations can either
%
\begin{inparaenum}
\item return an immediate subexpression of the same type as the input, or
\item swap the outermost data constructor with a (possibly) different one of the
  same type, reusing the immediate subexpressions of the input in any possible
  combination that produces a well-typed value.  
\end{inparaenum}

\setlength{\intextsep}{10pt}%
\setlength{\columnsep}{10pt}%
\begin{wrapfigure}{r}{0.2\textwidth}
\vspace{-20pt}
\begin{verbatim}
mutate t = case t of 
  Leaf x -> [ 
    PURE (Branch def x def)
  ]
  Branch l x r -> [ 
    PURE l, PURE r, 
    PURE (Leaf x), 
    PURE (Branch l x l),
    PURE (Branch r x r),
    PURE (Branch r x l)
  ]
\end{verbatim}
\vspace{-10pt}
\caption{\label{fig:mutagen:mutator}\mutagen mutator for the \texttt{Tree} data
  type.}
\vspace{-10pt}
\end{wrapfigure}

@@

% Defining mutations formally
Formally, for a data type $T$ defined in terms of the data constructors $C_i$,
each one with fields of (possibly) different types $t_j^i$:

\vspace{-10pt}
\begin{flalign*}
T \;:=\; C_1 \; t_1^1 \; t_2^1 \; \mydots  \;|\;
         C_2 \; t_1^2 \; t_2^2 \; \mydots  \;|\; 
         \mydots &&
\end{flalign*}
\vspace{-7pt}

\noindent Our tool automatically synthesizes \texttt{mutate} to follow the
equation:

\vspace{-10pt}
\begin{flalign*}
\texttt{mutate}(C_i \, x_1 \, x_2 \, \mydots) = 
  \mathtt{mut_{r}}(C_i \, x_1 \, x_2 \, \mydots) \cup
  \mathtt{mut_{s}}(C_i \, x_1 \, x_2 \, \mydots) && 
      % \setminus
      %  \{ (C_i \, x_1 \, x_2 \, \mydots) \} &&
\end{flalign*}
\vspace{-7pt}

\noindent Firstly, $\mathtt{mut_r}$ computes the set of possible mutations that
return an immediate subexpresion of the same type as the input:

\vspace{-10pt}
\begin{flalign*}
\mathtt{mut_{r}}(C_i \, x_1 \, x_2 \, \mydots) =  
  \{ 
    \mathtt{PURE}\, x_k
    \;|\;
    x_k \!\in\! \texttt{filter}(T, \{ x_1, \, x_2, \, \mydots \})
  \} &&
\end{flalign*}
\vspace{-7pt}

\noindent And $\mathtt{mut_s}$ builds all the possible mutations that swap the
root data constructor with a (possibly) different one, reusing (or defaulting)
compatible subexpresions to instantiate the new data constructor:

\vspace{-10pt}
\begin{flalign*}
\mathtt{mut_{s}}(C_i \, x_1 \, x_2 \, \mydots) \!=\! 
  \Vast\{ 
    \mathtt{PURE}(C_j \, x_1^\prime \, x_2^\prime \, \mydots)
    \,\Vast|\,
    \begin{aligned}
      & C_j^T \!\in\! \{ C_1, C_2, \mydots \} \\
      & x_k^\prime \!\in\! \overline{\mathtt{filter}}(t_k^j, \{x_1, \, x_2, \, \mydots \}) \\  
      & C_j \, x_1^\prime \, x_2^\prime \, \mydots \;\neq\; C_i \, x_1 \, x_2 \, \mydots
    \end{aligned}
  \Vast\} &&
\end{flalign*}
\vspace{-7pt}

\noindent Where the helper function $\mathtt{filter}$ simply returns the subset
of the input values $X$ that match a target type $t$:

\vspace{-10pt}
\begin{flalign*}
\mathtt{filter}(t, X) =  
  \{
    x
    \;|\;
    x \!\in\! X,
    \texttt{typeof}(x) = t
  \} &&
\end{flalign*}
\vspace{-7pt}

\noindent And the function $\overline{\mathtt{filter}}$ is a wrapper over
$\mathtt{filter}$ that returns the default value of the target type $t$ if there
are no compatible subexpressions available to reuse:

\vspace{-10pt}
\begin{flalign*}
\overline{\mathtt{filter}}(t, X) =
\Bigg\{
  \begin{aligned}
    & \mathtt{filter}(t, X) 
      &&\mathtt{if}\; \mathtt{filter}(t, X) \neq \emptyset \\
    & \{ \mathtt{def} \}
      &&\mathtt{otherwise}
  \end{aligned} &&
\end{flalign*}
\vspace{-7pt}

\noindent This ensures that a small constructor can always be grown into a
larger one by inserting default subexpresions whenever needed.
%
(Our previous example defined in Fig. \ref{fig:mutagen:mutator} show this in
practice for the case of mutating a \texttt{Leaf} into a \texttt{Branch}.)

% \begin{align*}
% \mathtt{mut_{r}}(C_i \; x_1 \; x_2 \; \mydots) = 
%   \{ \texttt{PURE}\, x_j \, | \, 
%   x_j \!\in\! \{ x_1, x_2, \mydots \}, \; 
%   \texttt{typeof}(x_j) = T \}
% \end{align*}

% \begin{align*}
% \mathtt{mut_{s}}(C_i \; x_1 \; x_2 \; \mydots) = 
%   \{ \texttt{PURE}\, (C_j \, x_1^\prime \, x_2^\prime \, \mydots) \, | \,
%   &C_j \!\in\! \{ C_1, C_2, \mydots \}, \, \\
%   &x_k^\prime \!\in\! \{ x_k \, | \, x_k \!\in\! \{ x_1, x_2, \mydots \}, \mathtt{typeof}(x_k) = t_k^j \}) \}
% \end{align*}

@@

%% At this point, the reader might consider:
%% %
%% what about large enumeration types like integers or characters?
%% %
%% Does \mutagen transform them exaustively too?
%% %
%% \emph{Certainly not.}
%
We can now focus on random mutants, which let us \emph{selectively avoid
exhaustiveness} when mutating values of large enumeration types (e.g. numbers).
  % 
  % --- or any
  % other type the user might want to use random mutations with.
%
%
Instead of creating a new \texttt{PURE} mutant for every other numerical
subexpression exhaustively, we condense them into a generator that can be
repeatedly sampled to produce new random values.   
%
% small number of values from a
% given random generator.
%
This way, a mutator for integers simply becomes:

@@

So far we have defined mutations that transform only the root of the input.
%
To apply these mutations to every internal subexpression we use two utility
functions.
% that can be automatically synthesized.
%
Firstly, a function \texttt{Positions} traverses the input and builds a Rose
tree\cite{meertens1988first} of \emph{mutable positions}, where positions are
lists of indices encoding the path from the root to every mutable subexpression.
%
% For instance, the positions of a \texttt{Tree} are computed as follows:
%
% \begin{verbatim}
% positions (Leaf x) =
%   node [ (0, positions x) ]
% positions (Branch l x r) =
%   node [ (0, positions l), (1, positions x), (2, positions r) ]
% \end{verbatim}
%
% \noindent where the function \texttt{node} simply builds a node of the positions
% Rose tree and propagates the accumulated prefix (the path from the root) to its
% children.
%
For instance, the mutable positions of the value \texttt{Branch (Leaf 1) 2
  (Leaf 3)} are as follows:

\vspace{-5pt}
\begin{equation}
  \nonumber
  \footnotesize
  \texttt{Positions}\quad
  \left(
  \tikz[
    baseline=-6mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\texttt{Branch}}
      child {node {\texttt{Leaf}}
        child {node {\texttt{1}}}
      }
      child {node {\texttt{2}}}
      child {node {\texttt{Leaf}}
        child {node {\texttt{3}}}
      };
  }
  \right)
  \quad
  =
  \quad
  \tikz[
    baseline=-6mm,
    level distance=0.5cm,
    level 1/.style={sibling distance=0.5cm},
    level 2/.style={sibling distance=0.5cm}
  ]{
    \node {\texttt{[]}}
      child {node {\texttt{[0]}}
        child {node {\texttt{[0,0]}}}
      }
      child {node {\texttt{[1]}}}
      child {node {\texttt{[2]}}
        child {node {\texttt{[2,0]}}}
      };
  }
\end{equation}
\vspace{-5pt}

\noindent Then, we define a function \texttt{MutateInside} that takes a desired
position within an input test case and mutates its corresponding subexpression,
returning a list of mutants.
%
%% given the desired position to mutate within a test case, we
%% define another function \texttt{inside} that finds the subexpression
%% corresponding to it and mutates it.
%
% A slightly simplified version of this function for the \texttt{Tree} data type is
% as follows:
%
% \begin{verbatim}
% inside []     t              = mutate t
% inside (0:ps) (Leaf x)       = [ Leaf x'       | x' <- inside ps x ]
% inside (0:ps) (Branch l x r) = [ Branch l' x r | l' <- inside ps l ]
% inside (1:ps) (Branch l x r) = [ Branch l x' r | x' <- inside ps x ]
% inside (2:ps) (Branch l x r) = [ Branch l x r' | r' <- inside ps r ]
% \end{verbatim}
%
This function traverses the desired position, calling itself recursively until
it reaches the desired subexpression, where a mutation encoded by
\texttt{mutate} can be applied directly.
%
The implementation of these two functions consists mostly of boilerplate code
that our tool automatically synthesizes, thus we avoid diving into details to
preserve space.

% locally at the top-level (case \texttt{inside [] t}).
%
% The rest of the \texttt{MutateInside}'s functionality simply unwraps and
% rewraps the intermediate subexpressions and is not particularly relevant for
% the point being made.

@@

% \subsection{Detecting Trace Space Saturation And Tuning Random Mutations Parameter}
\subsection{Tuning Random Mutations Parameter}

@@

\noindent Where $l_{pc}$, $l_{p}$, $l_{v}$, and $l_{v'}$ represent the labels
of: the program counter, the memory pointer, and the old and new values stored in
that memory cell.
%
The symbol $\vee$ simply denotes the join of two labels, i.e., the
\emph{maximum} of their sensibilities.
%
Later, bugs are systematically introduced in the IFC enforcing mechanism by
removing or weakening the checks stored in this rule table.
%
For instance, the following constitute the buggy variations of the
\texttt{Store} instruction:

%% \vspace{-10pt}
\begin{center}
\scriptsize
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Instruction} & \textbf{Precondition Check} & \textbf{Final PC Label} & \textbf{Final Result Label} \\
\hline
\texttt{Store} & $l_{pc} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
\texttt{Store} & $l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
\hline
\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $\bot$ & $l_{v'} \vee l_{pc} \vee l_{p}$ \\
\hline
\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{pc} \vee l_{p}$ \\
\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{p}$ \\
\texttt{Store} & $l_{pc} \vee l_{p} \sqsubseteq l_{v}$ & $l_{pc}$ & $l_{v'} \vee l_{pc}$ \\
\hline
\end{tabular}
\end{center}

@@

Having sorted these issues out, we mechanically injected 10 new bugs in the
validator as well as 5 new bugs in the interpreter of this engine (see Table
\ref{table:wasm:injectedbugs}).
%
These bugs either
%
\begin{inparaenum}
\item remove an existing integrity check (to weaken the WebAssembly
  type-system/validator); or
\item simulate a copy-paste bug \cite{CPMiner},%
  \footnote{This kind of bugs was inspired by the real bug \#1 we found prior to
    this step.}
  %
  replacing the implementation of an instruction with a compatible one
  (e.g., \texttt{i32.add} by \texttt{i32.sub}).
\end{inparaenum}

@@

%% We begin by designing a property to test the WebAssembly validator implemented
%% in \textit{haskell-wasm}.
%% %
Our approach is to assert that, whenever a randomly generated (or mutated)
WebAssembly module is valid according to \textit{haskell-wasm}, then the
reference implementation agrees upon it.
%
%% In other words, we are testing for false negatives.
%
In Haskell, we write the property:

@@

\noindent The precondition (\texttt{isValidHaskellWasm m}) runs the input
{Web-Assembly} module \texttt{m} against \textit{haskell-wasm}'s validator,
whereas the postcondition (\texttt{isValidRefImpl m}) serializes \texttt{m},
runs it against the reference implementation and checks that no errors are
produced.

@@

\begin{verbatim}
prop_interpreter ty fun args =
  do let m = mk_module fun ty "f"
     resHaskellWasm <- invokeHaskellWasm m "f" args
     resRefImpl     <- invokeRefImpl     m "f" args
     return (resHaskellWasm === resRefImpl)
\end{verbatim}

@@

%% If we only consider the succesfull runs where \fuzzchick does not time out, we
After carefully analyzing the results obtained using \fuzzchick, we observed a
peculiarity:
%
this tool either finds bugs quickly (after a few thousands tests) or does not
find them at all within the one-hour budget.
%% , suggesting that its power
%% scheduler assigns some enqueued test cases too little energy before they become
%% uninteresting and get ultimately thrown away.
%
With this in mind, it is fair to think that \fuzzchick might still be a better
choice if we set a shorter timeout and run it several times in a row.
%
Thus, to improve upon \fuzzchick, our tool must find bugs not only more
reliably, but also rapidly.

\setlength{\intextsep}{10pt}%
\setlength{\columnsep}{10pt}%
\begin{wrapfigure}{r}{0.5\linewidth}
%% \begin{wrapfigure}{r}{0.5\textwidth}
%% \centering
\vspace{-5pt}
\scriptsize

@@

\captionof{table}{\label{table:ifc:tests} Mean tests until first failure and effect size
  comparison between \fuzzchick and \mutagen when considering only successful
  runs.}
\vspace{-15pt}
%% \end{wrapfigure}
\end{wrapfigure}

Moreover, disabling the trace saturation heuristic (\textit{no reset}) we
observed two effects:
%
fixing the number of random mutations to 25 adds a constant overhead when
finding most of the (easier) bugs, suggesting that we are spending too time
mutating numeric subexpressions inside the generated machine states.
%
%% and that a
%% smaller number of random mutations could be equally effective to find such bugs.
%
However, some of the hardest-to-find bugs cannot be reliably exposed using this
fixed amount of random mutations (\#4, \#9, \#12, \#18, and \#10), suggesting
that one should increase this number even further to reliably discover all bugs
within the time budget.
%
In consequence, \emph{we argue that our heuristic is effective at automatically
tuning this parameter on the fly}.
%
%% , especially when the user is unsure about what the best value for it might be.

@@

% Limited evaluation
Our experiments cover generating test cases in two substantially different
scenarios (valid IFC machine states and WebAssembly programs), which are able to
find several planted and real previously-unknown bugs. 
%
Nonetheless, we cannot generalize that \mutagen will be effective at finding
bugs in other systems or testing properties. 
%
To compensate, our tool automatically synthesizes all the boilerplate code
required to operate using Template Haskell \cite{SheardJ02}, becoming an almost
drop-in replacement of \quickcheck (the most ubiquitous RPBT tool in Haskell). 
%     
This makes it an appealing alternative whenever \quickcheck is unable to
penetrate some sparse precondition and simply gives up.

% Coq -> Haskell translation
As mentioned in \S \ref{sec:casestudies}, our evaluation required us to
translate the IFC state machine used by \citeauthor{lampropoulos2019coverage}
from Coq to Haskell. 
%
Although this translation was performed mechanically (the process was partially
aided by Coq's code extraction mechanism), we cannot formally guarantee that our
Haskell version of the code behaves exactly as the original one. 
%

% However, we are highly confident that 
%
% In addition, one should also consider some  
% that the specific performance the
% runtime systems of these two programming languages can affect, the
% draw strong conclusions based only on time-to-failure.
